Echoes of Economic Downturn: Investigating the Persistent Impact of the Great Recession on Birth Rates Among Young Americans

### 2024-02-13 outputs/paper/paper.qmd
Finished variable table, removed todo for metho d and vriable, added explaination for variable as well.

# install.packages("mapdata")

library(tidyverse)
library(dplyr)
library(here)
library(mapdata)
library(ggplot2)
library(haven)
```

@@

## Methodology
// TODO: Add methodology for cleaning data

The data was processed and cleaned using R[@citeR], a powerful statistical programming language. Moreover, most of the data in our dataset are previously cleaned in the project we are replicating. Thus, figure 1 directly used data from `outputs/data/fig_1.csv` without any further cleaning necessary as it simply displays the birth rate from 1980 to 2020. Thus, the cleaned data is directly stored to `outputs/data/fig_1.csv` from `inputs/data/fig_1.csv`.

@@

In figure 3, the first dataset was imported from `inputs/data/fig_3a.csv` and the dataset containing American states information was from R[@mapdata]. The first dataset was first sliced to remove birth rate information regarding the states of Alaska and Hawaii as these two states can't be properly showned in the dataset from R[@mapdata]. Then, a temporary data frame is created with tibble from R[@tidyverse] witht the the abbreviated state names from `inputs/data/fig_3a.csv` and the state full names. Moreover, the abbreviated state name in the first dataset is switched with the state full names using left_join from R[@dpylr], which is then selected and renamed using R[@tidyverse]. After this, the adjusted dataset with the state full names and birth rate is merged with the second dataset that contains the American states information for further graphing.

@@

TODO: Add a table here to better understand the data.
```{r}
```

>>>>

# install.packages("mapdata")
# install.packages("kableExtra")

library(tidyverse)
library(dplyr)
library(here)
library(mapdata)
library(ggplot2)
library(haven)
library(kableExtra)
```

@@

## Methodology
The data was processed and cleaned using R[@citeR], a powerful statistical programming language. Moreover, most of the data in our dataset are previously cleaned in the project we are replicating. Thus, figure 1 directly used data from `outputs/data/fig_1.csv` without any further cleaning necessary as it simply displays the birth rate from 1980 to 2020. Thus, the cleaned data is directly stored to `outputs/data/fig_1.csv` from `inputs/data/fig_1.csv`.

@@

In figure 3, the first dataset was imported from `inputs/data/fig_3a.csv` and the dataset containing American states information was from R[@mapdata]. The first dataset was first sliced to remove birth rate information regarding the states of Alaska and Hawaii as these two states can't be properly showned in the dataset from R[@mapdata]. Then, a temporary data frame is created with tibble from R[@tidyverse] witht the the abbreviated state names from `inputs/data/fig_3a.csv` and the state full names. Moreover, the abbreviated state name in the first dataset is switched with the state full names using left_join from R[@dpylr], which is then selected and renamed using R[@tidyverse]. After this, the adjusted dataset with the state full names and birth rate is merged with the second dataset that contains the American states information for further graphing with R[@heaven].

@@

```{r}
#| label: tab-display-raw-data
#| tbl-cap: Birth rate of all groups from 1980 to 1989
#| echo: false

rawdata <- read.csv(here::here("inputs/data/figs_2a_2b.csv"))

rawdata |>
  select(year,brate_1519,brate_2024,brate_2529,brate_3034,brate_3539,brate_4044,brate_whitenh,brate_blacknh,brate_hisp) |>
  slice(1:10) |>
  kable(
    caption = "Birth rate of all groups from 1980 to 1989",
    col.names = c("Year", "15 to 19", "20 to 24","25 to 29", "30 to 34", "35 to 39", "40 to 44", "white", "black", "hispanic"),
    digits = 1,
    booktabs = TRUE
  )
```

This table presents the birth rates for various age groups and racial categories in the United States from the year 1980 to 1989. Moreover, the birth rates are typically expressed per every 1,000 women in the population.To further understand this table, Year represent the years from the selected year range. The next six columns are different age groups for women, which are 15 to 19 years, 20 to 24 years, 25 to 29 years, 30 to 34 years, 35 to 39 years, and 40 to 44 years. Moreover, the last three columns are ethnic groups that describes the birth rates among the white, black, and Hispanic populations. However, some data for these groups are not available except 1989.
